---
title: "Trading Analysis in Turkey"
subtitle: "Rsızlar"
institute: "Mef University"
date: "2020/12/26"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      
---

class: inverse, center, middle

# Table of Contents

* Our Goals
* About Trading Dataset
* Line Graph
* Mapping
* Reporting Part
* Thanks for Listening

---

class: inverse, center, middle

# Goals

Our goal is to achieve logical results by using monthly import and export data between 2013 and 2020 by visualizing and analyzing their distribution by cities for Turkey. In addition, we aimed to reach the balance between export and import figures.
---

class: inverse, center, middle

# About Dataset

In order to create the date variable in the dataset we used, we created values from 1 to 12 for the months in the "Month" variable in the dataset and created a new date format. Thus, we were able to reach the increase and decrease in Import and Export by date. 
Since we encountered a Turkish character problem, we fixed the errors in these parts.

---

class: inverse, center, middle

# Line Graph

When we visualized the import and export figures by months of the last two years, we observed that the difference was at least in November 2018 and at most in August 2020.
While the period with the highest import figures is September 2020, the period with the highest export figures is May 2019.

---

class: inverse, center, middle

```{r eval=TRUE, fig.width=10.5, fig.height=6,echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(devtools)
library(leaflet)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(rio)
library(leaflet)
library(shiny)
library(treemap)
library(plotly)
library(ggmap)
library(leaflet.extras)
githubURL <- "https://github.com/pjournal/mef04g-rsizlar/blob/gh-pages/Data/data.RData?raw=true"
load(url(githubURL))

#Clean Month and City variables
data$Month<-str_sub(data$Month,start= 4)

data<-data %>% 
  separate(City, c("CityCode", "City"),sep = "( - )")

data<-data%>%
  mutate(Month=case_when(data$Month=="Ocak"~"01",
                         data$Month=="Subat"~"02",
                         data$Month=="Mart"~"03",
                         data$Month=="Nisan"~"04",
                         data$Month=="Mayis"~"05",
                         data$Month=="Haziran"~"06",
                         data$Month=="Temmuz"~"07",
                         data$Month=="Agustos"~"08",
                         data$Month=="Eylul"~"09",
                         data$Month=="Ekim"~"10",
                         data$Month=="Kasim"~"11",
                         data$Month=="Aralik"~"12"))

data<-data%>%
  mutate(TradingType=case_when(data$TradingType=="Ithalat"~"Import",
                               data$TradingType=="Ihracat"~"Export"))

#Create Date Variable from Year and Month
data$Date<-paste0(data$Year,as.character("-"),data$Month,as.character("-01"))
data$Date<-as.Date(data$Date, format="%Y-%m-%d")

data<-left_join(data,reg)

turkceden_ingilizceye <- function(dataset){
  turkce_harfler<- c("Ç","Ş","Ğ","İ","Ü","Ö","ç","ş","ğ","ı","ü","ö")
  ingilizce_harfler<- c("C","S","G","I","U","O","c","s","g","i","u","o")
  dataset=mgsub(turkce_harfler,ingilizce_harfler,dataset)
  return(dataset)
}
  

mgsub <- function(pattern, replacement, x, ...) {
  n = length(pattern)
  if (n != length(replacement)) {
    stop("pattern and replacement do not have the same length.")
  }
  result = x
  for (i in 1:n) {
    result <- gsub(pattern[i],replacement[i],result)
  }
  return(result)
}

tur@data$NAME_1 <- turkceden_ingilizceye(tur@data$NAME_1 )
tur@data$NAME_1 <- gsub("K. Maras", "Kahramanmaras",tur@data$NAME_1 )
tur@data$NAME_1 <- gsub("Kinkkale","Kirikkale",tur@data$NAME_1 )
tur@data$NAME_1 <- gsub("Zinguldak", "Zonguldak", tur@data$NAME_1 )
tur@data$NAME_1 <- gsub("Afyon","Afyonkarahisar", tur@data$NAME_1 )

tur@data$NAME_1 <- toupper(tur@data$NAME_1) 

data <-data %>% left_join(koordinat,by=c("City"))   

data_mean <- data  %>% 
  filter(Year>=2018)%>%
  select(City,TradingType,AmountUSD,LONGITUDE,LATITUDE)%>%
  group_by(City,TradingType,LONGITUDE,LATITUDE) %>% 
  summarise(AmountUSD=sum(AmountUSD))%>%
  pivot_wider(names_from = TradingType, values_from = AmountUSD)%>%
  mutate(Balance=round((Export-Import)/1000000,2))%>%
  select(City,LONGITUDE,LATITUDE,Balance)

tur@data <- tur@data%>%left_join(data_mean,by=c("NAME_1"="City"))

cols<- c("Export"="#F8766D","Import"="#00BFC4","Balance"="#FFDB6D")

p<-data%>%
  filter(Date>"2018-08-01")%>%
  select(Date,TradingType,AmountUSD)%>%
  group_by(Date,TradingType)%>%
  summarise(Total=round(sum(AmountUSD)/1000000000,2))%>%
  pivot_wider(names_from = TradingType, values_from = Total)%>%
  mutate(Balance = Export-Import)%>%
  ggplot() +
  geom_col(aes(x=Date, y=Balance, fill = "Balance"),size = 0.5, color = "#C4961A")+
  geom_line(aes(x=Date, y=Export,color="Export"),size=1) +
  geom_line(aes(x=Date, y=Import,color="Import"),size=1) +
  scale_colour_manual(name="Lines",values=cols) + scale_fill_manual(name="Bar",values=cols)+
  theme_test() + 
  labs(x="Date", y="Total Amount (billion dollars)", title = "Yearly Total Export/Import Comparison")

ggplotly(p,tooltip = c("Balance","Export","Import","Date"))%>%
  layout(legend = list(orientation = "h", x = 0.25, y = -0.2))

```

---
class: inverse, center, middle

# Mapping

We used the map method in order to examine the Import and Export figures according to the provinces in more detail.
In this way, by clicking on the cities, you can better understand the distribution between Export and Import over the years and access its details.

---

class: inverse, center, middle
```{r eval=TRUE, fig.width=10,5, fig.height=6,echo=FALSE, message=FALSE, warning=FALSE}


labels <- sprintf("<strong>%s</strong><br/>%s $ ",tur@data$NAME_1,prettyNum(tur@data$Balance, format = "f", big.mark = ".",decimal.mark ="," )
) %>% lapply(htmltools::HTML)

bins <- c(-Inf,-15000, -5000, 0, 1000, 3000, 5000, 10000)
pal <- colorBin("viridis", domain =tur@data$Balance, bins = bins)

leaflet(tur)  %>% 
  addProviderTiles("Stamen.TonerBackground", group = "Terrain")%>%
  addProviderTiles("CartoDB.DarkMatterNoLabels", group = "CartoDB"  ) %>% 
  addProviderTiles("Esri.OceanBasemap", group = "Esri") %>% 
  addPolygons(smoothFactor = 0.3, fillOpacity = 10,weight = 2,
              opacity = 1,color = "white",
              dashArray = "3",
              highlight = highlightOptions(stroke = 2,
                                           weight = 4,
                                           color = "#666",
                                           dashArray = NULL,
                                           fillOpacity = 0.7,
                                           bringToFront = TRUE),fillColor = ~pal(Balance),label=labels,labelOptions = labelOptions(
                                             style = list("font-weight" = "normal", padding = "3px 8px"),
                                             textsize = "15px",
                                             direction = "auto")) %>%
  addLegend(pal = pal, values = ~Balance, opacity = 1.0) %>%
  
  addLayersControl(baseGroups = c( "CartoDB", "Esri","Terrain")) %>% 
  setView(lng = 33.243322, lat =	41.963745, zoom =5.4)%>% addSearchOSM()%>% addReverseSearchOSM()%>% addResetMapButton()

```

---
class: inverse, center, middle

# Reporting Parts

When we examined the Import and Export figures by years, we observed that the Marmara region is at the highest level every year.

---

class: inverse, center, middle
```{r eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE,fig.align="center"}
total_data<-data%>%
    filter(Year==2019)%>%
    select(City,TradingType,AmountUSD)%>%
    group_by(City,TradingType)%>%
    summarise(Amount = sum(AmountUSD))%>%
    pivot_wider(names_from = TradingType, values_from = Amount,values_fill = 0)

table_export<-left_join(total_data,pop)%>%
    select(City,Export,Pop,MenPop,WomenPop)%>%
    mutate(ExportPP = round(Export/Pop,2),
           WomenRatio = round(WomenPop/(WomenPop+MenPop),2))%>%
    arrange(desc(ExportPP))

table_export
```
Above table gives 2019 Trading figures sorted by Export amount per Person for each city. First 5 cities have huge industry operations and their Export metrics are better than other cities.
---
class: inverse, center, middle

# Thanks for Listening
